\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-09-04 smurf_paper.tex
some more work on SMURF description

As described in the introduction, the approach taken by SMURF to
reducing \scuba\ data is to model (and remove) noise sources, and to
iterate this process along with estimates of the map. This sort of
algorithm has been used successfully in the past with instruments
including SCUBA \citep{1999ASPC..172..171J}, SHARC-2
\citep{kovacs2008}, and BOLOCAM \citep{aguirre2010}, and is capable of
recovering modest angular scales up to of order the footprint of the
array (depending, of course, on the noise in the data). The main
benefit of this approach over CMB-style maximum-likelihood methods
(which are generally superior at recovering structure on all scales
available in the data) is that the computing resources are
significantly smaller, both in terms of memory usage and execution
time. Given the massive volume of \scuba\ data produced in a single
observation (a rate of approximately 10 GB\,hr$^{-1}$ in each focal
plane), iterative techniques were the only viable option for providing
real-time feedback at the telescope, and for automatically reducing
near-science grade maps on a timescale comensurate with the
observation times using standard high-end desktop computers.  We do
note, however, that recent advances in direct inversion methods such
as SANEPIC \citep{patanchon2008} may in fact be tenable for \scuba\
data. A future goal of the software team is to investigate meshing
SMURF (perhaps as a cleaning step and initial condition) with a
SANEPIC-like algorithm.

Significant experimentation with different models for noise sources
during \scuba\ commissioning required a highly flexible software
framework with a simple (though extensive) interface for user
configurability. Though written in C, \smurf\ uses an object-oriented
style of handling the

>>>>

As described in the introduction, the approach taken by SMURF to
reducing \scuba\ data is to model (and remove) predominantly
low-frequency noise sources that are correlated amongst detectors, and
to iterate this process along with estimates of the map. This sort of
algorithm has been used successfully in the past with instruments
including SCUBA \citep{1999ASPC..172..171J}, SHARC-2
\citep{kovacs2008}, and BOLOCAM \citep{aguirre2010}, and is capable of
recovering modest angular scales up to of order the footprint of the
array (depending, of course, on the noise in the data). The main
benefit of this approach over CMB-style maximum-likelihood methods
(which are generally superior at recovering structure on all scales
available in the data) is that the computing resources are
significantly smaller, both in terms of memory usage and execution
time. Given the massive volume of \scuba\ data produced in a single
observation (a rate of approximately 10 GB\,hr$^{-1}$ in each focal
plane), iterative techniques were the only viable option for providing
real-time feedback at the telescope, and for automatically reducing
near-science grade maps on a timescale comensurate with the
observation times using standard high-end desktop computers.  We do
note, however, that recent advances in direct inversion methods such
as SANEPIC \citep{patanchon2008} may in fact be tenable for \scuba\
data. A future goal of the software team is to investigate meshing
SMURF (perhaps as a cleaning step and initial condition) with a
SANEPIC-like algorithm.

Significant experimentation with different models for noise sources
during \scuba\ commissioning required a highly flexible software
framework with a simple interface for user configurability. To achieve
this goal, while minimizing development time, it was decided to create
SMURF as a Starlink package \citep{2009ASPC..411..418J}, which
provides access to a large suite of libraries (including a commanding
and messaging interface, astrometric conversions and generation of
standard WCS information, file formats etc.). Furthermore, Starlink is
already used extensively at the Joint Astronomy Centre (host of the
JCMT) for many other systems, which helps with
interoperability. Though originally written primarily in FORTRAN, many
of the core Starlink libraries are now ported into native C, or else
have, at least, a C interface. It was therefore decided to develop
SMURF in C (rather than C++ which would have required adding further
dependencies to Starlink), although the noise models are implemented
with an object-oriented approach (all data associated with models are
stored in structures, and there is a standard interface for method
functions). In this way it is fairly easy to add new
models. Parallelisation is incorporated in the most time-consuming
low-level routines (e.g. performing Fast Fourier Transforms,
re-gridding the data), usually handling either independent blocks of
bolometers, or blocks of time, in each thread, depending on the nature
of the calculation. ... something about scalability ...
